Text Is All You Need

### 2023-05-05 docs/hes_final/article.qmd

In this work we explored the possibility of predicting future geopolitical conflict escalation at a country level using a global collection of news publications. We compared the usage of text-only data and data enriched with country labels and social indicators. We experimented with different ways of aggregating the data and applied both statistical machine learning and deep learning models. The results suggest that text embeddings provide a strong enough signal to make such predictions. Different data sampling and modeling approaches produced comparable results and each of the proposed models performed best at some of the common numeric metrics. The neural network-based models appear to generalize better across geographies while statistical model provided a size/cost of deployment advantage.

>>>>

In this work we explored the possibility of predicting future geopolitical conflict escalation at a country level using a global collection of news publications. We compared the usage of text-only data and data enriched with country labels and social indicators. We experimented with different ways of aggregating the data and applied both statistical machine learning and deep learning models. The results suggest that text embeddings provide a strong enough signal to make such predictions. Different data sampling and modeling approaches produced comparable results and each of the proposed models performed best at some of the common numeric metrics. The neural network-based models appear to generalize better across geographies while the statistical model provided a size/cost of deployment advantage.
